---
title: "R Markdown and Leaflet"
author: "Iaroslav Iefimenko"
date: '14-01-2018'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r echo=FALSE, message=FALSE, results = 'hide'}
#all the packages that we will use 
library(dplyr)
library(leaflet)
```

## Introduction

This project was created as part of the Developing Data Products course of the Coursera Data Science Specialisation. The goal of the project is to create a web page using R Markdown wth a map created with Leaflet, and to host the resulting web page.

## Airports

[OurAirports](http://ourairports.com/about.html) is a free site where visitors can explore the world's airports, read other people's comments, and leave their own. For current example we use a CSV-formatted data dump of all our airports, countries, and regions, which is updated every night. 

```{r echo=FALSE}
airports <- read.csv2("http://ourairports.com/data/airports.csv",header=TRUE, sep=",", encoding = "UTF-8", stringsAsFactors = FALSE)
```
Structure of airports data:
```{r echo=FALSE}
str(airports)
airports$latitude_deg <- as.numeric(airports$latitude_deg)
airports$longitude_deg <- as.numeric(airports$longitude_deg)
```

## Airports map building
Build custom icons:
```{r}
planeIcons <- icons(
  iconUrl = ifelse(airports$type == "heliport", "https://iaroslav-iefimenko.github.io/RandLeaflet/icons/helicopter-airport.png",
    ifelse(airports$type == "small_airport", "https://iaroslav-iefimenko.github.io/RandLeaflet/icons/small-airport.png", 
    ifelse(airports$type == "closed", "https://iaroslav-iefimenko.github.io/RandLeaflet/icons/closed-airport.png",
    ifelse(airports$type == "seaplane_base", "https://iaroslav-iefimenko.github.io/RandLeaflet/icons/sea-plane-airport.png",
    ifelse(airports$type == "balloonport", "https://iaroslav-iefimenko.github.io/RandLeaflet/icons/ballon-airport.png",
    ifelse(airports$type == "medium_airport", "https://iaroslav-iefimenko.github.io/RandLeaflet/icons/medium-airport.png",
    ifelse(airports$type == "large_airport", "https://iaroslav-iefimenko.github.io/RandLeaflet/icons/large-airport.png",
    "https://iaroslav-iefimenko.github.io/RandLeaflet/icons/unknown-airport.png"))))))),
  iconWidth = 16, iconHeight = 16,
  iconAnchorX = 16, iconAnchorY = 16
)
```

Build and show map:
```{r}
map <- leaflet(data = airports) %>% addTiles() %>%
  addMarkers(lng = ~longitude_deg, lat = ~latitude_deg, label = ~name,
    clusterOptions = markerClusterOptions(), icon = planeIcons,
    popup = ~paste(sep = "<br/>",
    "<b>Airport name:</b>", name, "<b>Airport type:</b>", type, "<b>ISO Country:</b>", iso_country,
    "<b>Ident:</b>", ident, "<b>Home link:</b>", home_link, "<b>Wikipedia link:</b>", wikipedia_link
))
map
```